\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{sectsty}

\geometry{margin=0.65in}
\setlist[itemize]{left=0pt, itemsep=0pt}
\subsubsectionfont{\normalfont}

\begin{document}
\pagestyle{empty}

\begin{center}
  \section*{\LARGE Ronald David Martinez Marenco}
  \href{mailto:ronaldavidmm2006@gmail.com}{ronaldavidmm2006@gmail.com} |
  \href{https://github.com/RonaldDavidMartinezMarenco/}{https://github.com/RonaldDavidMartinezMarenco/} |
  \href{https://www.linkedin.com/in/ronald-m/}{https://www.linkedin.com/in/ronald-m/}

  Barranquilla, Colombia
\end{center}

\hrule
\vspace{4pt}

% Professional Summary
\subsection*{Summary}
Software Engineering Student (3rd year) passionate about the tech world, artificial intelligence, 
and eventually becoming an innovator. with experience in Python, Machine Learning, and backend
development through academic and personal projects. Developed AI-driven prototypes incluiding chatbots
and computer vision systems using YOLO and Tensorflow. Familiar with building REST APIs using FastAPI,
and creating basic web interfaces with JavaScript, HTML, CSS, and React. Strong academic background and focused 
on growing skills in applied AI and data-driven systems.

% Work Experience
\subsection*{Projects}

\subsubsection*{%
  \textbf{Data Viento: Weather Climate Insights Platform} - \href{https://dataviento.voltr.org//}{DataViento Webpage}
  \hfill
  \textit{October 2025 - Present}
}
Python, FastAPI, JavaScript, HTML, CSS, MySQL, Docker, OpenMeteo API, Cron Automation, Gemini LLM
\begin{itemize}

  \item Developed a full-stack web application that collects, stores, 
        and visualizes climate data from the OpenMeteo API, providing 
        interactive dashboards and AI-assisted explanations based on the 
        user's selected city.
  \item Built backend services using FastAPI to consume weather data from
        the OpenMeteo API and store structures time-series data in MySQL,
        ensuring efficient retrieval and analysis.
  \item Implemented frontend intercafe with HTML, CSS,JavaScript and chart.js library to render a
        dashboard that displays weather variables and trend charts  based on the user's 
        selected city.
  \item Developed an Gemini LLM-powered chatbot to explain weather variables, chart
        interpretations, and climate impacts applying prompt engineering to 
        improve contextual responses.
  \item Automated data ingestion via Python Scripts and Linux cron jobs, updating
        database on a schedule without manual intervetion.
  \item Docker environment for containerized deployment and Git version control to
        manage the full project lifecycle collaboratively.

\end{itemize}

\subsubsection*{%
  \textbf{Hunting Exoplanets NASA-Hackathon} - \href{https://exoplanet-app-730407208895.us-central1.run.app/}{Hunting Exoplanet Webpage}
  \hfill
  \textit{October 2025}
}
Python, Tensorflow, Google Colab, Streamlit.
\begin{itemize}
  \item Developed a predictive web application for exoplanet classification,
         allowing users to input planetary features or upload CSV files for batch 
         prediction using an interactive Streamlit interface.
  \item Cleaned and preprocessed astronomical datasets in Google Colab, handling missing 
        values using median imputation and performing correlation analysis to understand 
        feature relationships.
  \item Built and evaluated multiple ensemble machine learning models (Voting Classifier combining RandomForest, LGBM, XGBoost) within a 
        TensorFlow-compatible architecture to ensure robust prediction performance.
  \item Achieved approximately 87\% accuracy on the trained dataset.
      
\end{itemize}
\subsubsection*{%
  \textbf{Educational AI Chatbot with RAG and Voice Interface} - \href{https://github.com/RonaldDavidMartinezMarenco/Mini-Chatbot}{Github Repository}
  \hfill
  \textit{April 2025 - June 2025}
}
Java, Spring Boot, Spring AI, MySQL, HTML, Tailwind CSS, JavaScript, Gemini (Vertex AI), ElevenLabs, Jina AI Embeddings
\begin{itemize}
  \item Developed an educational chatbot backend using Spring Boot and Spring AI, 
        supporting conversational memory and intent-aware responses.
  \item Integrated generative responses via Gemini (Vertex AI) and added text-to-speech (TTS)
        functionality using ElevenLabs for voice interaction.
  \item Implemented Retrieval-Augmented Generation (RAG) by extracting text from educational
        documents using Apache Tika and generating embeddings with Jina AI, enabling document-based 
        question answering.
  \item Built a lightweight frontend with Tailwind CSS and persisted conversation and metadata in a 
        MySQL database.\\
\end{itemize}

\subsubsection*{%
  \textbf{Colombian License Recognition} - \href{https://github.com/RonaldDavidMartinezMarenco/license_recognition}{Github Repository}
  \hfill
  \textit{January 2025 - February 2025}
}
Python, YOLOv11n, OpenCV, EasyOCR, Computer Vision, CSV Data Pipelines
\begin{itemize}
  \item Developed a real-time license plate recognition system for Colombian 
        plates using YOLOv11n to detect plates from images,videos, and live camera streams.
  \item Integrated EasyOCR to extract license plate text as structured strings after image preprocessing.
  \item Built an image processing pipeline with OpenCV, applying grayscale conversion, normalization, 
        and standardization to improve OCR accuracy.
  \item Designed a real-time data pipeline to capture detected plates, handle missing values, and 
        append results dynamically to a CSV file.
  \item Achieved approximately 98\% accuracy on the trained dataset through iterative training and evaluation.

\end{itemize}
\subsubsection*{%
  \textbf{Fundacion Remedios de Manases Webpage (Freelance Work)} - \href{https://fundacion-remedios.vercel.app/}{Landing Page}
  \hfill
  \textit{November 2024}
}
React, JavaScript, HTML, CSS (Responsive Design)
\begin{itemize}
  \item Developed a responsive and dynamic website based on client requirements for clarity, 
        accessibility, and lightweight design.
  \item Implemented dynamic UI behavior using plain CSS, applying subtle hover transitions 
        (scale and rotation effects) to enhance interactivity without overloading the interface.
  \item Aligned the interface with the foundation's branding, ensuring consistency with logo, colors, 
        and typography.
  \item Designed a fully responsive layout, ensuring readability and usability across desktop and mobile
        devices through custom CSS breakpoints.
\end{itemize}
\subsubsection*{%
  \textbf{Sign Language Recognition App Using Machine Learning} - \href{https://bonga.unisimon.edu.co/server/api/core/bitstreams/83c0cc48-143b-40a7-b3c8-34bda57d3a5e/content}{Scientific article}
  \hfill
  \textit{March 2024 - Present}
}
Python, Tensorflow, Keras, Mediapipe Holistic, OpenCV, Numpy, LSTM / CNN Neural Networks
\begin{itemize}
  \item Developed a sign language recognition application using computer vision and deep 
        learning to classify 10 Colombian sign gestures.
  \item Created a custom dataset by recording and labeling sign language samples, then expanded
        it using image and sequence augmentation techniques.
  \item Extracted pose, hand, and facial landmarks with MediaPipe Holistic and standardized
        input data before model training.
  \item Trained an LSTM-based neural network using TensorFlow and Keras to model temporal 
        gesture sequences.
  \item Achieved approximately 95\% classification accuracy on the trained gesture set through iterative training and evaluation.

\end{itemize}

% Education
\subsection*{Education}
\subsubsection*{\textbf{B.Sc. in Systems Engineering} from Universidad Simon Bolivar, Colombia \hfill \textit{In Progress - 5th Semester
}}
\begin{itemize}
  \item 1st Place in GPA - Systems Engineering Program \\
        GPA: 4.86 / 5.0
  \item Active member of the Data Science research group.

  \item Student representative for the systems engineering program.
\end{itemize}

%Honor and Awards
\subsection*{Honor \& Awards}
\begin{itemize}
  \item \textbf{NASA Space Apps Challenge - Global Nominee} \\
        \textbf{Project}: Hunting Exoplanets - Developed an AI-based solution for detection of exoplanets using Python, Tensorflow, Pandas, Streamlit. 
  \item \textbf{Elite Researcher Recognition}\\
        Awarded for academic research excellence.
  \item \textbf{RedColsi Speaker at international conference}\\
        Speaker for \textbf{Sign Language Recognition App Using Machine Learning}
        
  
  
\end{itemize}
% Skills
\subsection*{Skills}
\begin{itemize}
  \item \textbf{Languages:} Python, JavaScript, SQL.
  \item \textbf{Web:} HTML, CSS, React.
  \item \textbf{Backend:} FastAPI, Flask.
  \item \textbf{AI/Machine Learning:} Tensorflow, YOLO, NumPy, Pandas.
  \item \textbf{Technologies:} MySQL, Docker, Git, Github.
\end{itemize}

% TODO: Projects section
\end{document}
